---
title: "Vitamins"
output: html_notebook
---

Setup

```{r}
rm(list=ls()) #clear workspace


library(tidyverse)
library(reshape2)
library(ggplot2)
library(ggtree)
library(ggtreeExtra)
library(ggnewscale)

```



##Vitamin auxotrophy screen

Load data
```{r}
data <- readRDS("data/2021-02-time_course.rds")

#We focused on 72 hour timepoint. Remove the two later timepoints, and two preculture timepoints 

data <- data %>% subset(!timept=="120hrs") %>% 
  subset(!timept=="96hrs") %>% 
  subset(!timept=="preculture48hr") %>% 
  subset(!timept=="preculture24hr")
```

Generate a single value to represent sensitivity:

median(with-vit OD)-median(without-vit OD)/(with-vit OD + without-vit OD)

```{r}

data_subset <- subset(data, timept== "72hrs")

#Summarize based on median OD values

dataSummary <- data_subset %>%
  group_by(vit, strains) %>%
  summarize(median= median(value, na.rm=TRUE), min = min(value), max= max(value), IQR =IQR(value))

colnames(dataSummary)

##cast 
data_simple <- dcast(dataSummary, strains~vit, value.var = "median")

rownames(data_simple) <- data_simple$strains
data_simple <- data_simple[2:3]

#Normalize and calculate the difference between median growth with and without vitamins
data_simple <- data_simple/rowSums(data_simple)
data_simple$diff <- data_simple$minus-data_simple$plus

```

Fig. S2. Visualize distribution of values to find appropriate cutoff. 
```{r}
hist(as.matrix(data_simple$diff), breaks=40, main=paste("Change in growth with and without vitamins"), xlab=paste("Δ growth (norm.)"))

##choose cut off of 0.4

##These 52 isolates were selected for further screening to verify auxotrophies
sensitive <- subset(data_simple, diff < -0.4)

```

##Results of growth assays to verify auxotrophies 
Growth of each suspected auxotroph with and without each vitamin individually. 

```{r}

panel <- readRDS("data/vitpanels_batch5678_screen.rds")
##simplify your life and only take the second timepoint for visualization
#data_full <- data


panel <- panel %>% 
  separate(vitamin, c("vitamin_simple",NA), remove=F)

panel <- subset(panel, timept=="72hrs")

#Example of panel data: A2R20, as in Fig 1c

panel_subset <- panel %>% subset(strains=="A2R20")
  
ggplot(panel_subset, aes(x=vitamin, y=value, color=type)) +
  facet_wrap(strains~., ncol=4, scales= "fixed")+
  theme_bw(base_size = 15) + 
  stat_summary(fun = median,
                 geom = "crossbar", width = 0.8)+
  geom_point(shape=21, size=1, fill="white")+
  labs(x= "", y="OD 600")+
  scale_color_manual(values=c("#fb9a99", "#1f78b4"))+
  theme(axis.text.x = element_text(angle = 90, hjust=1))+
  ggtitle("")+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(strip.background =element_rect(fill="white",linewidth=0))


```

##Fig. 1c inset

Combine data from screen and for B1 from panel for Fig. 1c inset. 
```{r}

data_A2R20 <- panel %>% 
    subset(strains=="A2R20") %>%
    subset(vitamin_simple=="B1") %>%
  subset(vitamin !="B1_thi-pyro")

##remove the B1_thi knockout
data_A2R20_added <- data_A2R20 %>% subset(vitamin =="B1_thiamine" & type=="added") %>%
  subset(select=c("strains","timept","type","value"))

data_A2R20_out <- data_A2R20 %>% subset(vitamin =="B1_both") %>%
  subset(select=c("strains","timept","type","value"))

data_A2R20_added$condition <- "B1 added"
data_A2R20_out$condition <- "B1 removed"

data_subset <- subset(data, strains=="A2R20")

data_subset <- subset(data_subset, select=c("strains","timept","vit","value"))

colnames(data_subset) <- c("strains", "timept", "type", "value")

data_subset$condition <- data_subset$timept

data_A2R20 <- rbind(data_subset,data_A2R20_added, data_A2R20_out)


ggplot(data_A2R20, aes(y=value, x=condition, color=type)) +
  facet_wrap(strains~., scales="fixed", ncol=4)+
  theme_bw(base_size = 10) + 
  stat_summary(fun = median,
                 geom = "crossbar", width = 0.8)+
  geom_point(shape=21, size=1, fill="white")+
  labs(x= "OD 600", y="")+
  scale_color_manual(values=c("black", "black","#fb9a99", "#1f78b4"))+
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
  theme(strip.background =element_rect(fill="white",linewidth=0))

ggsave("figures/A2R20.svg", device="svg", width=4, height=2)

```

##Visualize phylogenetic tree- Fig 1c


```{r}
#Results of panel summarized in metadata file
metadata_sensitive <- as.data.frame(read.csv("data/metadata_sensitive_strains.csv"))

##import phylogenetic tree
tree <- read.tree("data/prunedGenomes.tree")

# import phylogeny of full isolate collection
phylo <- read.csv("data/taxonomy-151strains.csv", row.names = 1)

##For visualization, melt auxotrophies
sens.melt <- melt(metadata_sensitive,measure.vars = c("auxo_1", "auxo_2", "auxo_3", "auxo_4"))

sens.melt <-sens.melt[!sens.melt$value=="", ]

```



Visualize tree with auxotrophies

```{r}

colors <- c("#80C87F", #altero
            "brown", #bacill
            "orange", #cyto
             "#FEFE99", #flavo
            "darkgreen", #gran
             "#5CB7FF", #pseudo
              "#FFCFFD", #rhodo
              "#BEADD4") #vibrio

##first make the tree skeleton
p <- ggtree(tree) 

p1 <- p + 
      geom_tiplab(
        size=2.5) +
      geom_fruit(
          data=phylo,
          geom=geom_point,
          mapping=aes(y=taxa, color=order),
          position="identity"
          ) +   scale_color_manual(values = colors)


p2 <- p1+
  new_scale_fill() +
   geom_fruit(
          data=sens.melt,
          geom=geom_tile,
          mapping=aes(y=strains, x=value, fill=value),
          offset=0.07,   # The distance between layers, default is 0.03 of x range of tree.
          pwidth=0.15
          ) 
p2


```
save image

```{r}
ggsave("figures/tree-screen-heatmap.svg", p2, svg, scale=2, height=5)
```









